A unified model for parsing and sentence understanding

### 2016-03-14 writing/hard_stack_paper/paper.tex
Initial tweaks and responses to JG paper edits.

\newcommand{\shift}{\textsc{push}}
\newcommand{\reduce}{\textsc{merge}}

@@

The resulting vector $\vec{p}_{trans}$ is a probability distribution over parsing decisions (i.e., predicting whether to \shift or \reduce at the current timestep). In the full PINN model, we follow whichever transition has a higher predicted probability in this distribution. This prediction module is trained to mimic the decisions of an external parser. In particular, we use the transition sequences corresponding to the constituency parses included in the SNLI corpus. The classifier is trained using a cross-entropy objective:
\begin{equation}
  L_{trans} = - \log \vec{p}_{trans}[t^*]
\end{equation}
where $t^*$ is the correct transition specified by the gold parse, and $\vec{p}[\cdot]$ represents vector indexing.

@@

A naive implementation of the PINN model would require simulating a stack of size $N$ for each input sentence. In order to update the model weights via backpropagation, we would also need to maintain all intermediate stack representations for later chain rule calculations. This implies a total space requirement of $M \times N \times T \times D$, which is prohibitively large for significant batch sizes $M$ or sentence lengths $N$. Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift and \reduce operations only affect the top of the stack).

We designed an alternative minimal-space stack algorithm inspired by the zipper technique \cite{huet1997zipper}. For a single input sentence, we represent the stack over the entire feedforward with a $T \times D$ matrix $S$. Each row $S_t$ represents the top of the actual stack at timestep $t$. We maintain a queue of backpointers onto $S$ in order to track which elements should be involved in a \reduce~operation at any given time. Algorithm~\ref{alg:thin-stack} below describes the full mechanics of a stack feedforward in this compressed representation. It operates on the compressed $T \times D$ matrix $S$ and a backpointer queue $Q$. Table~\ref{tbl:thin-stack} shows an example run of this algorithm.

@@

\begin{table}
\centering
\begin{tabular}{c|cl}
  \toprule
  $t$ & $S$[$t$] & $Q_t$ \\
  \midrule
  0 & $a$ & 0 \\
  1 & $b$ & 0 1 \\
  2 & $c$ & 0 1 2 \\
  3 & $M(c, b)$ & 0 3 \\
  4 & $M(M(c, b), a)$ & 4 \\
  \bottomrule
\end{tabular}
\caption{An example of the thin-stack algorithm computing a \shift-\shift-\shift-\reduce-\reduce sequence on the input sentence $(a, b, c)$. $S$ is shown in the second column and may be thought of as a list of the tops of the stack at all timesteps $t$. The last two elements of $Q$ specify the rows $t$ which should be involved in a \reduce~operation at the next timestep.}
\label{tbl:thin-stack}
\end{table}

@@

@@

\begin{table*}[t]
  \centering\small
  \begin{tabular}{lcccc} 
    \toprule
Model                   & Params.    & Trans. acc.  &   Train  &   Test \\
\midrule
\multicolumn{5}{c}{Previous non-NN results}\\
\midrule
Lexicalized classifier \cite{snli:emnlp2015}
                        & --                & --                    &   99.7   &   78.2      \\
\midrule
\multicolumn{5}{c}{Previous sentence encoder-based NN results}\\
\midrule
100d LSTM encoders \cite{snli:emnlp2015}
                        & 221k               & --               &   84.8   &   77.6      \\
1024d pretrained GRU encoders \cite{DBLP:journals/corr/VendrovKFU15}
                        & 15m                & --              &   98.8   &   81.4       \\
300d Tree-based CNN encoders \cite{mou2015recognizing}
                        & 3.5m                & --             &   83.4   &   82.1       \\
\midrule
\multicolumn{5}{c}{Our results}\\
\midrule
300d LSTM RNN encoders          & 3.0m                  & --                &   83.9      &   80.6       \\
\result{300d PINN-Ext-NoTracking}   
                        & ?                  & --                &   ?      &   81.9dv       \\
\result{300d PINN-Ext }
                        & ?                  & --                &   ?      &   \underline{84.8dv}       \\
\result{300d PINN }
                        & ?                  & 92.2+dv            &   ?    &   83.3+dv       \\          
% \result{300d PINN-Ext, sequence-based attn.   }     
%                         & ?                  & --                &   ?      &   86.3dv         \\
% \result{300d PINN-Ext, tree-based attn. }           
%                         & ?                  & --                &   ?      &   \textbf{?}\\
\midrule
\multicolumn{5}{c}{Other previous NN results}\\
\midrule
100d LSTM w/ word-by-word attention \cite{rocktaschel2015reasoning}
                        & 252k               & --              &   85.3   &   83.5       \\
300d mLSTM word-by-word attention model \cite{DBLP:journals/corr/WangJ15b}
                        & 1.9m               & --             &   92.0   &   86.1      \\
300d LSTMN with deep attention fusion \cite{cheng2016long}
                        & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
    \bottomrule
  \end{tabular}
  \protect\caption{\protect\label{tab:results}Results on SNLI 3-way inference classification. Params. is the approximate number of trained parameters (excluding word embeddings for models where they are trained). Trans. acc. is the model's accuracy in predicting parsing transitions. Train and test are SNLI classification accuracy. \todo{[SB]: Estimate param counts as numbers come in.}} 
\end{table*}

>>>>

\newcommand{\shift}{\textsc{shift}}
\newcommand{\reduce}{\textsc{reduce}}

@@

The resulting vector $\vec{p}_{trans}$ is a probability distribution over parsing decisions (i.e., predicting whether to \shift~or \reduce~at the current timestep). In the full PINN model, we follow whichever transition has a higher predicted probability in this distribution. This prediction module is trained to mimic the decisions of an external parser. In particular, we use the transition sequences corresponding to the constituency parses included in the SNLI corpus. The classifier is trained using a cross-entropy objective:
\begin{equation}
  L_{trans} = - \log \vec{p}_{trans}[t^*]
\end{equation}
where $t^*$ is the correct transition specified by the gold parse, and $\vec{p}[\cdot]$ represents vector indexing.

@@

A naive implementation of the PINN model would require simulating a stack of size $N$ for each input sentence. In order to update the model weights via backpropagation, we would also need to maintain all intermediate stack representations for later chain rule calculations. This implies a total space requirement of $B \times N \times T \times D$, which is prohibitively large for significant batch sizes $B$ or sentence lengths $N$. Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift~and \reduce~operations only affect the top of the stack).

We designed an alternative minimal-space stack algorithm inspired by the zipper technique \cite{huet1997zipper}. For a single input sentence, we represent the stack with a single $T \times D$ matrix $S$. \todo{[JG] Either expand or remove references to batch size.} Each row $S_t$ represents the top of the actual stack at timestep $t$. We maintain a queue of backpointers onto $S$ in order to track which elements should be involved in a \reduce~operation at any given time. Algorithm~\ref{alg:thin-stack} below describes the full mechanics of a stack feedforward in this compressed representation. It operates on the compressed $T \times D$ matrix $S$ and a backpointer queue $Q$. Table~\ref{tbl:thin-stack} shows an example run of this algorithm.

@@

\begin{table}
\centering
\begin{tabular}{c|cl}
  \toprule
  $t$ & $S$[$t$] & $Q_t$ \\
  \midrule
  0 & $a$ & 0 \\
  1 & $b$ & 0 1 \\
  2 & $c$ & 0 1 2 \\
  3 & $(c~b)$ & 0 3 \\
  4 & $((c~b)~a)$ & 4 \\
  \bottomrule
\end{tabular}
\caption{An example of the thin-stack algorithm computing a \shift-\shift-\shift-\reduce-\reduce~sequence on the input sentence $(a, b, c)$. $S$ is shown in the second column and may be thought of as a list of the tops of the stack at all timesteps $t$. The last two elements of $Q$ specify the rows $t$ which should be involved in a \reduce~operation at the next timestep.}
\label{tbl:thin-stack}
\end{table}

@@

\begin{table*}[t]
  \centering\small
  \begin{tabular}{lcccc} 
    \toprule
Model                   & Params.    & Trans. acc.  &   Train  &   Test \\
\midrule
\multicolumn{5}{c}{Previous non-NN results}\\
\midrule
Lexicalized classifier \cite{snli:emnlp2015}
                        & --                & --                    &   99.7   &   78.2      \\
\midrule
\multicolumn{5}{c}{Previous sentence encoder-based NN results}\\
\midrule
100d LSTM encoders \cite{snli:emnlp2015}
                        & 221k               & --               &   84.8   &   77.6      \\
1024d pretrained GRU encoders \cite{DBLP:journals/corr/VendrovKFU15}
                        & 15m                & --              &   98.8   &   81.4       \\
300d Tree-based CNN encoders \cite{mou2015recognizing}
                        & 3.5m                & --             &   83.4   &   82.1       \\
\midrule
\multicolumn{5}{c}{Our results}\\
\midrule
300d LSTM RNN encoders          & 3.0m                  & --                &   83.9      &   80.6       \\
\result{300d PINN-Ext-NoTracking}   
                        & 3.4m                  & --                &   ?      &   81.9dv       \\
\result{300d PINN-Ext }
                        & 3.7m                  & --                &   ?      &   \underline{84.8dv}       \\
\result{300d PINN }
                        & ??                  & 92.2+dv            &   ?    &   83.3+dv       \\          
% \result{300d PINN-Ext, sequence-based attn.   }     
%                         & ?                  & --                &   ?      &   86.3dv         \\
% \result{300d PINN-Ext, tree-based attn. }           
%                         & ?                  & --                &   ?      &   \textbf{?}\\
\midrule
\multicolumn{5}{c}{Other previous NN results}\\
\midrule
100d LSTM w/ word-by-word attention \cite{rocktaschel2015reasoning}
                        & 252k               & --              &   85.3   &   83.5       \\
300d mLSTM word-by-word attention model \cite{DBLP:journals/corr/WangJ15b}
                        & 1.9m               & --             &   92.0   &   86.1      \\
300d LSTMN with deep attention fusion \cite{cheng2016long}
                        & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
    \bottomrule
  \end{tabular}
  \protect\caption{\protect\label{tab:results}Results on SNLI 3-way inference classification. Params. is the approximate number of trained parameters (excluding word embeddings for models where they are trained). Trans. acc. is the model's accuracy in predicting parsing transitions. Train and test are SNLI classification accuracy.} 
\end{table*}

@@
